### 2022-12-06 dissertation/appendix/symbols.tex
✏️ (final-review): References, datasets, start of symbols.

@@

\todo[inline]{ref usage}

\begin{description}
    \item
    [\parbox{1cm}{$\alpha$}] Control parameter for leaky ReLU that controls non-zero gradients.

    \item
    [\parbox{1cm}{$\beta$}] Batch size.

    \item
    [\parbox{1cm}{$c$}] Index for classes.

    \item
    [\parbox{1cm}{$C$}] Total number of classes.

    \item
    [\parbox{1cm}{$E$}] Total number of epochs.

    \item
    [\parbox{1cm}{$\epsilon$}] Error value calculated from loss/cost function.

    \item
    [\parbox{1cm}{$f$}] The activation function implemented by the \ac{AN}.

    \item
    [\parbox{1cm}{$f_{AN}$}] The non-linear function represented by the \ac{AN}.

    \item
    [\parbox{1cm}{$\vec{h}$}] The vector representing the activations of the hidden units.

    \item
    [\parbox{1cm}{$h_j$}] The $j$-th hidden unit and thus its activation.

    \item
    [\parbox{1cm}{$i$}] Index for the input dimensions.

    \item
    [\parbox{1cm}{$I$}] Total number of dimensions on the input space.

    \item
    [\parbox{1cm}{$j$}] Index for the hidden layer.

    \item
    [\parbox{1cm}{$J$}] The total number of units in the hidden layer.

    \item
    [\parbox{1cm}{$k$}] Index for the output dimensions.

    \item
    [\parbox{1cm}{$K$}] The total number of dimensions in the output space.

    \item
    [\parbox{1cm}{$\lambda$}] Control parameter for the steepness of sigmoid and hyperbolic tangent activation functions.

    \item
    [\parbox{1cm}{$net$}] The net input signal.

    \item
    [\parbox{1cm}{$net'$}] The augmented net input signal which contains the bias term $\theta$.

    \item
    [\parbox{1cm}{$\mu_i$}] The mean over all input patterns at index $i$.

    \item
    [\parbox{1cm}{$p$}] Index used for patterns in the set of all training data.

    \item
    [\parbox{1cm}{$P$}] Total number of patterns in the training data.

    \item
    [\parbox{1cm}{$\omega_{min}$}] The lower bound of the range for random uniform sampling.

    \item
    [\parbox{1cm}{$\omega_{max}$}] The upper bound of the range for random uniform sampling.

    \item
    [\parbox{1cm}{$\mathbb{R}$}] The real number space.

    \item
    [\parbox{1cm}{$\sigma$}] The standard deviation.

    \item
    [\parbox{1cm}{$\sigma_i^2$}] Unit variance.

    \item
    [\parbox{1cm}{$T$}] The total number of dimensions in the target space.

    \item
    [\parbox{1cm}{$\theta$}] The bias or threshold term.

    \item
    [\parbox{1cm}{$\vec{v}$}] The vector of weights associated with the input.

    \item
    [\parbox{1cm}{$v_i$}] The $i$-th index of the weight vector associated with the input.

    \item
    [\parbox{1cm}{$v_{i,j}$}] The $i,j$-th index of the weight vector between the input and hidden layers.

    \item
    [\parbox{1cm}{$\vec{w}$}] The weight vector between the hidden and output layers.

    \item
    [\parbox{1cm}{$w_{j,k}$}] The $j,k$-th index of the weight vector between the hidden and output layers.

    \item
    [\parbox{1cm}{$\vec{x}$}] The input vector.

    \item
    [\parbox{1cm}{$\vec{x}_p$}] The input vector for pattern $p$.

    \item
    [\parbox{1cm}{$x_i$}] The $i$-th index of the input vector $\vec{x}$

    \item
    [\parbox{1cm}{$x_{i,p}$}] The $i$-th index of the input vector $\vec{x}$ for pattern $p$.

    \item
    [\parbox{1cm}{$x_{i,p}'$}] The pre-processed input at the $i$-th index of the input vector $\vec{x}$ for pattern $p$.

    \item
    [\parbox{1cm}{$x_{i_{min}}$}] The minimum value at the  $i$-th index of the input vector $\vec{x}$  for all patterns in $P$.

    \item
    [\parbox{1cm}{$x_{i_{max}}$}] The maximum value at the  $i$-th index of the input vector $\vec{x}$  for all patterns in $P$.

    \item
    [\parbox{1cm}{$\vec{y}$}] The output vector.

    \item
    [\parbox{1cm}{$\vec{y}_p$}] The output vector for pattern $p$.

    \item
    [\parbox{1cm}{$y_k$}] The $k$-th dimension of the output vector $\vec{y}$.

    \item
    [\parbox{1cm}{$y_{k}'$}] The post-processed output at the $k$-th dimension of the output vector $\vec{y}$.

    \item
    [\parbox{1cm}{$y_{k,p}$}] The $k$-th dimension of the output vector $\vec{y}$ for pattern $p$.

    \item
    [\parbox{1cm}{$y_{k,p}'$}] The post-processed output at the $k$-th dimension of the output vector $\vec{y}$ for pattern $p$.

    \item
    [\parbox{1cm}{$\hat{y}_{k,p}$}] The $k$-th dimension of the target output for pattern $p$. 
\end{description}

>>>>

\section{Chapter~\ref{chap:introduction}: Introduction}
\label{sec:symbols:introduction}

@@

\begin{description}
	\item
	      [\parbox{1cm}{$\alpha$}] Control parameter for leaky ReLU that controls non-zero gradients. (See Eq.~\eqref{eq:an_function_mapping} p.~\pageref{eq:an_function_mapping})
\end{description}

\section{Chapter~\ref{chap:heuristics}: Heuristics}
\label{sec:symbols:heuristics}

\section{Chapter~\ref{chap:hhs}: Hyper-Heuristics}
\label{sec:symbols:hhs}

\section{Chapter~\ref{chap:probability}: Probability}
\label{sec:symbols:probability}

\section{Chapter~\ref{chap:bhh}: Bayesian Hyper-Heuristic}
\label{sec:symbols:bhh}

\section{Chapter~\ref{chap:methodology}: Methodology}
\label{sec:symbols:methodology}

\section{Chapter~\ref{chap:results}: Results}
\label{sec:symbols:results}

\section{Chapter~\ref{chap:conclusion}: Conclusion}
\label{sec:symbols:conclusion}

### 2022-12-06 dissertation/chapters/01-introduction.tex
✏️ (final-review): References, datasets, start of symbols.

\Acf{ML} is one of the most popular fields of research in \acf{AI} studies today. In recent years, \acs{ML} research has seen some notable achievements in academia~\cite{ref:lecun:2015, ref:glorot:2010, ref:goodfellow:2014, ref:quoc:2017}, as well as the industry at large~\cite{ref:silver:2016, ref:silver:2017, ref:zoph:2017, ref:lewis:2017}.  \acs{ML} research has grown tremendously over the past decade with successes like AlphaGo, which set new standards for \acs{AI} capabilities by beating the world's best Go player, Lee Sedol, 4-1~\cite{ref:san-hun:2016}.

@@

Techniques exist to dynamically balance the trade-off between exploration and exploitation during the search process. An example of such a technique is to dynamically adjust and learn the heuristic \index{hyper-parameters}\textit{hyper-parameters} as part of the learning process. This field of study is known as \index{meta-learning}meta-learning~\cite{ref:giraud:2004}. \index{meta-learning}Meta-learning of heuristic hyper-parameters as applied in the training of \acp{ANN} has shown to yield good generalisation results~\cite{ref:hospedales:2020, ref:vilalta:2002}.

@@

\begin{itemize}
      \item \textbf{Appendix~\ref{app:datasets}} provides details on the datasets used for the empirical analysis.

@@

      \item \textbf{Appendix~\ref{app:acronyms}} provides a list of the important acronyms used or newly defined in the course of this work, as well as their associated definitions.

>>>>

\Acf{ML} is one of the most popular fields of research in \acf{AI} studies today. In recent years, \acs{ML} research has seen some notable achievements in academia~\cite{ref:lecun:2015, ref:glorot:2010, ref:goodfellow:2014, ref:vaswani:2017}, as well as the industry at large~\cite{ref:silver:2016, ref:silver2017, ref:zoph:2017, ref:lewis:2017}.  \acs{ML} research has grown tremendously over the past decade with successes like AlphaGo, which set new standards for \acs{AI} capabilities by beating the world's best Go player, Lee Sedol, 4-1~\cite{ref:silver:2017}.

@@

Techniques exist to dynamically balance the trade-off between exploration and exploitation during the search process. An example of such a technique is to dynamically adjust and learn the heuristic \index{hyper-parameters}\textit{hyper-parameters} as part of the learning process. This field of study is known as \index{meta-learning}meta-learning~\cite{ref:giraud:2004}. \index{meta-learning}Meta-learning of heuristic hyper-parameters as applied in the training of \acp{ANN} has shown to yield good generalisation results~\cite{ref:hospedales:2022, ref:vilalta:2002}.

@@

\begin{itemize}
      \item \textbf{Appendix~\ref{app:acronyms}} provides a list of the important acronyms used or newly defined in the course of this work, as well as their associated definitions.

      \item \textbf{Appendix~\ref{app:symbols}} lists and defines the mathematical symbols used in this work, categorised according to the relevant chapter in which they appear.

      \item \textbf{Appendix~\ref{app:datasets}} provides details on the datasets used for the empirical analysis.

@@

### 2022-12-06 dissertation/chapters/02-artificial-neural-networks.tex
✏️ (final-review): References, datasets, start of symbols.

The biological neural systems are made up of microscopic nerve cells called neurons~\cite{ref:jain:1996}. Figure~\ref{fig:biological_neuron} illustrates a single \acs{BN}. The main components of the \acs{BN} include the cell body, \index{dendrites}\textit{dendrites} and the \index{axon}\textit{axon}. \Acp{NN} are formed through the connections between the \index{axon}axons and \index{dendrites}dendrites of various neurons. This is known as \index{synaptogenesis}\textit{synaptogenesis}~\cite{ref:huttenlocher:1997}. Such a connection is referred to as a \index{synapse}\textit{synapse}. Communication takes place, through the synapse, by electro-chemical pulse and is often referred to as an \textit{activation} or \textit{action potential}.  Communication signals propagate from the \index{dendrites}dendrites, through the cell body to the axon of a neuron, provoking a signal in the post-synaptic neuron~\cite{ref:engelbrecht:2007}. The greater the connection between two neurons, the stronger the communication.  Kennedy~\cite{ref:kennedy:2016} defines stronger \index{synapse}synapses as ones that contribute more depolarisation to the neural membrane upon activation than weaker ones. Stronger \index{synapse}synapses have a higher probability of generating an action potential in the post-synaptic neuron. During activation, the pre-synaptic neuron release neurotransmitters that bind to the post-synaptic neuron~\cite{ref:khanacademy:synapse}. The frequent release of these molecules cause the \index{synapse}synapse to grow. Connections that grow over time yield stronger signals (learning), while connections that are weak propagate low intensity signals and vanish over time (forgetting). The ability of synapses to strengthen and weaken over time is known as \index{synaptic plasticity}\textit{synaptic plasticity}~\cite{ref:huttenlocher:1997}.

@@

For qualitative data, the class labels have to be converted from textual to numerical representations. In general, class labels are encoded as numerical vectors~\cite{ref:srinidhi:2018, ref:brownlee:2017:one-hot}. One such encoding technique is referred to as \index{one-hot encoding}\textit{one-hot encoding}. Harris and Harris~\cite{ref:harris:2010} describe \index{one-hot encoding}one-hot encoding as a group of vector elements where the classes are represented with a single high element (represented by $1$) and all the others low (represented by $0$). Each element $y_k$, where $k$ is the $k$-th element of the encoded vector $\boldsymbol{y}$, uniquely refers to a class. An activation, represented by a value of $1$ at a given index thus represents a class associated with that index.

@@

Weight initialisation is dependent on the \index{activation function}activation function used. If weights are initialised as values that are too small, the vanishing gradients problem can occur~\cite{ref:hanin:2018}. If weights are initialised as values that are too big, output of the activation function would not be in the active range. This leads to unit saturation, and exploding gradients can occur~\cite{ref:hanin:2018, ref:yadav:2018}.

Many different weight initialisation techniques have been developed~\cite{ref:erdogmus:2003}. Some techniques have been developed to address the problems that are mentioned above~\cite{ref:yadav:2018}. For the purposes of this dissertation, focus is put on \index{random uniform sampling}random uniform sampling, \index{Glorot uniform sampling}\textit{Glorot uniform (Xavier)} sampling and \index{Glorot normal sampling}\textit{Glorot normal} sampling~\cite{ref:glorot:2010}. Brief discussions on each of these weight initialisation techniques are presented as follows.

@@

An advantage of \acs{ReLU} is that it is not susceptible to the \index{vanishing gradients problem}vanishing gradients problem~\cite{ref:xu:2015, ref:maksutov:2018} which occurs when gradients in the first layers of a multi-layer \acs{ANN} approach zero, and have no effect on the training process.

@@

The \acf{LReLU} \index{activation function}activation function is a variant of the \acs{ReLU} \index{activation function}activation function that avoids zero gradients in the negative part of its argument by introducing a scaling parameter, $\alpha > 0$~\cite{ref:xu:2015}. Similar to \acs{ReLU}, \acs{LReLU} is not susceptible to the \index{vanishing gradients problem}vanishing gradients problem. However, \acs{LReLU} does not suffer from the \index{dying ReLU problem}dying \acs{ReLU} problem~\cite{ref:agarap:2018}, which occurs when the neurons that use \acs{ReLU} \index{activation function}activation functions become \textit{inactive} and only output $0$ due to a negative net input signal. The \acs{LReLU} \index{activation function}activation function is given as

>>>>

The biological neural systems are made up of microscopic nerve cells called neurons~\cite{ref:jain:1996}. Figure~\ref{fig:biological_neuron} illustrates a single \acs{BN}. The main components of the \acs{BN} include the cell body, \index{dendrites}\textit{dendrites} and the \index{axon}\textit{axon}. \Acp{NN} are formed through the connections between the \index{axon}axons and \index{dendrites}dendrites of various neurons. This is known as \index{synaptogenesis}\textit{synaptogenesis}~\cite{ref:huttenlocher:1997}. Such a connection is referred to as a \index{synapse}\textit{synapse}. Communication takes place, through the synapse, by electro-chemical pulse and is often referred to as an \textit{activation} or \textit{action potential}.  Communication signals propagate from the \index{dendrites}dendrites, through the cell body to the axon of a neuron, provoking a signal in the post-synaptic neuron~\cite{ref:engelbrecht:2007}. The greater the connection between two neurons, the stronger the communication. Kennedy~\cite{ref:kennedy:2016} defines stronger \index{synapse}synapses as ones that contribute more depolarisation to the neural membrane upon activation than weaker ones. Stronger \index{synapse}synapses have a higher probability of generating an action potential in the post-synaptic neuron. During activation, the pre-synaptic neuron release neurotransmitters that bind to the post-synaptic neuron. The frequent release of these molecules cause the \index{synapse}synapse to grow. Connections that grow over time yield stronger signals (learning), while connections that are weak propagate low intensity signals and vanish over time (forgetting). The ability of synapses to strengthen and weaken over time is known as \index{synaptic plasticity}\textit{synaptic plasticity}~\cite{ref:huttenlocher:1997}.

@@

For qualitative data, the class labels have to be converted from textual to numerical representations. In general, class labels are encoded as numerical vectors~\cite{ref:brownlee:2017:one-hot}. One such encoding technique is referred to as \index{one-hot encoding}\textit{one-hot encoding}. Harris and Harris~\cite{ref:harris:2010} describe \index{one-hot encoding}one-hot encoding as a group of vector elements where the classes are represented with a single high element (represented by $1$) and all the others low (represented by $0$). Each element $y_k$, where $k$ is the $k$-th element of the encoded vector $\boldsymbol{y}$, uniquely refers to a class. An activation, represented by a value of $1$ at a given index thus represents a class associated with that index.

@@

Weight initialisation is dependent on the \index{activation function}activation function used. If weights are initialised as values that are too small, the vanishing gradients problem can occur~\cite{ref:hanin:2018}. If weights are initialised as values that are too big, output of the activation function would not be in the active range. This leads to unit saturation, and exploding gradients can occur~\cite{ref:hanin:2018}.

Many different weight initialisation techniques have been developed~\cite{ref:erdogmus:2003}. For the purposes of this dissertation, focus is put on \index{random uniform sampling}random uniform sampling, \index{Glorot uniform sampling}\textit{Glorot uniform (Xavier)} sampling and \index{Glorot normal sampling}\textit{Glorot normal} sampling~\cite{ref:glorot:2010}. Brief discussions on each of these weight initialisation techniques are presented as follows.

@@

An advantage of \acs{ReLU} is that it is not susceptible to the \index{vanishing gradients problem}vanishing gradients problem~\cite{ref:xu:2015} which occurs when gradients in the first layers of a multi-layer \acs{ANN} approach zero, and have no effect on the training process.

@@

The \acf{LReLU} \index{activation function}activation function is a variant of the \acs{ReLU} \index{activation function}activation function that avoids zero gradients in the negative part of its argument by introducing a scaling parameter, $\alpha > 0$~\cite{ref:xu:2015}. Similar to \acs{ReLU}, \acs{LReLU} is not susceptible to the \index{vanishing gradients problem}vanishing gradients problem. However, \acs{LReLU} does not suffer from the \index{dying ReLU problem}dying \acs{ReLU} problem~\cite{ref:trottier:2017}, which occurs when the neurons that use \acs{ReLU} \index{activation function}activation functions become \textit{inactive} and only output $0$ due to a negative net input signal. The \acs{LReLU} \index{activation function}activation function is given as

### 2022-12-06 dissertation/chapters/03-heuristics.tex
✏️ (final-review): References, datasets, start of symbols.

\Acf{NAG} is a variant of the \acs{Momentum} heuristic developed by~\citeauthor{ref:sutskever:2013-2}~\cite{ref:sutskever:2013-2} and is inspired by~\citeauthor{ref:nesterov:1983}'s~\cite{ref:nesterov:1983} work on optimising convex functions. \acs{NAG} provides an improvement to the momentum accumulation term by providing a look-ahead term that better refines the weight update step.

In the \acs{NAG} heuristic, the gradient is not calculated w.r.t the current weights, but rather w.r.t the approximate future positions of the weights~\cite{ref:sutskever:2013-2, ref:ruder:2016}. An illustration of the weight update vector using \acs{NAG} is taken from Geoffrey Hinton's lecture on mini-batch \acs{GD}~\cite{ref:hinton:2012} and is presented in Figure~\ref{fig:heuristics:gd:nag} below.

@@

The difference in the update step for \acs{Momentum} and the update step for \acs{NAG} as developed by~\citeauthor{ref:sutskever:2013-2}~\cite{ref:sutskever:2013-2} is described by~\citeauthor{ref:ruder:2016}~\cite{ref:ruder:2016} as follows. \acs{Momentum} first calculates the current gradient, represented by the small \textcolor{blue}{blue} vector in Figure~\ref{fig:heuristics:gd:nag} and then takes a large step in the direction of the updated accumulated gradient, presented by the big \textcolor{blue}{blue} vector. In contrast, \acs{NAG} first takes a big step in the direction of the previous accumulated gradient presented by the \textcolor{brown}{brown} vector. At this point the gradient is measured and \acs{NAG} makes a correction represented by the \textcolor{red}{red} vector. The complete update step is represented by the green vector. By anticipating approximate future positions of the weights, the weight update step as defined by \acs{NAG} controls the optimisation process from going too fast and results in increased responsiveness~\cite{ref:bengio:2013}. Along with the same velocity update step for \acs{Momentum}, as presented in Equation~\eqref{eq:heuristics:gd:momentum_part_1}, the \acs{NAG} weight update step is given as

>>>>

\Acf{NAG} is a variant of the \acs{Momentum} heuristic developed by~\citeauthor{ref:sutskever:2013}~\cite{ref:sutskever:2013} and is inspired by~\citeauthor{ref:nesterov:1983}'s~\cite{ref:nesterov:1983} work on optimising convex functions. \acs{NAG} provides an improvement to the momentum accumulation term by providing a look-ahead term that better refines the weight update step.

In the \acs{NAG} heuristic, the gradient is not calculated w.r.t the current weights, but rather w.r.t the approximate future positions of the weights~\cite{ref:sutskever:2013, ref:ruder:2016}. An illustration of the weight update vector using \acs{NAG} is taken from Geoffrey Hinton's lecture on mini-batch \acs{GD}~\cite{ref:hinton:2012} and is presented in Figure~\ref{fig:heuristics:gd:nag} below.

@@

The difference in the update step for \acs{Momentum} and the update step for \acs{NAG} as developed by~\citeauthor{ref:sutskever:2013}~\cite{ref:sutskever:2013} is described by~\citeauthor{ref:ruder:2016}~\cite{ref:ruder:2016} as follows. \acs{Momentum} first calculates the current gradient, represented by the small \textcolor{blue}{blue} vector in Figure~\ref{fig:heuristics:gd:nag} and then takes a large step in the direction of the updated accumulated gradient, presented by the big \textcolor{blue}{blue} vector. In contrast, \acs{NAG} first takes a big step in the direction of the previous accumulated gradient presented by the \textcolor{brown}{brown} vector. At this point the gradient is measured and \acs{NAG} makes a correction represented by the \textcolor{red}{red} vector. The complete update step is represented by the green vector. By anticipating approximate future positions of the weights, the weight update step as defined by \acs{NAG} controls the optimisation process from going too fast and results in increased responsiveness~\cite{ref:bengio:2013}. Along with the same velocity update step for \acs{Momentum}, as presented in Equation~\eqref{eq:heuristics:gd:momentum_part_1}, the \acs{NAG} weight update step is given as

### 2022-12-06 dissertation/chapters/05-probability.tex
✏️ (final-review): References, datasets, start of symbols.

The difference between the frequentist approach and the Bayesian approach can be illustrated using an example.~\citeauthor{ref:hackenberger:2019}~\cite{ref:hackenberger:2019} suggests an experiment that investigates whether the gender ratio in some hypothetical mice population is $1:1$. Two experiments were designed. In the first experiment, mice are randomly selected until the first male is chosen. The result in this experiment will then be the total number of mice chosen by gender. For the second experiment, exactly seven mice are randomly selected. The result of the second experiment would be the number of males and females in a sample of seven. Suppose the outcome of the experiment was $FFFFFFM$, where $F$ represents a female and $M$ represents a male. If the experimental design is not known ahead of time, the result is useless. Consider the $p$-value for each of these experiments. The $p$-value is the probability of obtaining results at least as extreme as the observed results of a statistical hypothesis test~\cite{ref:beers:2022}. For the first experiment, the $p$-value is $0.031$ and for the second experiment, the $p$-value is $0.227$. Using a confidence level of $\alpha = 0.05$, opposite outcomes could be concluded for these two experiments when it comes to rejecting the null hypothesis, despite using the same data. The reason for the difference in outcomes, is due to the difference in their null distributions, which represent the probability distribution of the test statistic when the null hypothesis is true. The first approach uses a geometrical approach, and the second used a binomial approach as illustrated in Figure \ref{fig:probability:bayesian_statistics:mouse_experiment_outcome}.

>>>>

The difference between the frequentist approach and the Bayesian approach can be illustrated using an example.~\citeauthor{ref:hackenberger:2019}~\cite{ref:hackenberger:2019} suggests an experiment that investigates whether the gender ratio in some hypothetical mice population is $1:1$. Two experiments were designed. In the first experiment, mice are randomly selected until the first male is chosen. The result in this experiment will then be the total number of mice chosen by gender. For the second experiment, exactly seven mice are randomly selected. The result of the second experiment would be the number of males and females in a sample of seven. Suppose the outcome of the experiment was $FFFFFFM$, where $F$ represents a female and $M$ represents a male. If the experimental design is not known ahead of time, the result is useless. Consider the $p$-value for each of these experiments. The $p$-value is the probability of obtaining results at least as extreme as the observed results of a statistical hypothesis test~\cite{ref:thisted:1998}. For the first experiment, the $p$-value is $0.031$ and for the second experiment, the $p$-value is $0.227$. Using a confidence level of $\alpha = 0.05$, opposite outcomes could be concluded for these two experiments when it comes to rejecting the null hypothesis, despite using the same data. The reason for the difference in outcomes, is due to the difference in their null distributions, which represent the probability distribution of the test statistic when the null hypothesis is true. The first approach uses a geometrical approach, and the second used a binomial approach as illustrated in Figure \ref{fig:probability:bayesian_statistics:mouse_experiment_outcome}.
